---
title: 'Peer Graded Assignment: PML'
author: "Jerry Odia"
date: "January 16, 2019"
output: html_document
---
Overview
The document is the final report of the Peer Assessment project from Coursera's course Practical Machine Learning in data science specializtion. 
The goal of the project is to predict the manner in which 6 participants performed some exercise as described below. This is the "classe" variable in the training set. The machine learning algorithm described here is applied to the 20 test cases available in the test data and the predictions are submitted in appropriate format to the Course Project Prediction Quiz for automated grading.

Environment Preparation
```{r (echo = TRUE)
```
```{r}
library(knitr)
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
library(corrplot)
library(nlme)
set.seed(33245)
```
Data loading and Cleaning
```{r}
UrlTrain <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
UrlTest  <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
#download the datasets
training <- read.csv(url(UrlTrain))
testing  <- read.csv(url(UrlTest))
#create data partion
inTrain  <- createDataPartition(training$classe, p=0.7, list=FALSE)
TrainSet <- training[inTrain, ]
TestSet  <- training[-inTrain, ]
```
```{r}
dim(TrainSet)
dim(TestSet)
```
```{r}
# remove variables that are mostly NA
AllNA    <- sapply(TrainSet, function(x) mean(is.na(x))) > 0.95
TrainSet <- TrainSet[, AllNA==FALSE]
TestSet  <- TestSet[, AllNA==FALSE]
dim(TrainSet)
```
```{r}
# remove Near Zero variables
NZV <- nearZeroVar(TrainSet)
TrainSet <- TrainSet[, -NZV]
TestSet  <- TestSet[, -NZV]
dim(TrainSet)
```
```{r}
TrainSet<-TrainSet[,-c(2,5)]
TestSet<-TestSet[,-c(2,5)]
```
Standardize by preProcessing 
```{r}
preobj<-preProcess(TrainSet,method=c("center","scale"))
TrainSetS<-predict(preobj,TrainSet)
TestSetS<-predict(preobj,TestSet)
```
Correlation Analysis
```{r}
corMatrix<-cor(TrainSetS[,-57])
corrplot(corMatrix,order="FPC",method="color",type = "lower",tl.cex=0.8, tl.col = rgb(0,0,0))
```
Prediction Modelling
Three methods(models) will be trained using the TrainSets, the one with highest accuracy when applied to the TestSets will be used for the quiz predictions. The methods are: Random Forests, Decision Tree and Generalized Boosted Model, as described below.
A Confusion Matrix is plotted at the end of each analysis to better visualize the accuracy of the models.
```{r}
# Random Forest model fit
set.seed(33245)
controlRF <- trainControl(method="cv", number=3, verboseIter=FALSE)
modFitRF <- train(classe ~ ., data=TrainSetS, method="rf",trControl=controlRF)
modFitRF$finalModel                          
```
```{r}
# prediction on Test dataset
predictRF <- predict(modFitRF, newdata=TestSetS)
confMatRF <- confusionMatrix(predictRF, TestSetS$classe)
confMatRF
```
```{r}
#plot matrix results
plot(confMatRF$table, col = confMatRF$byClass, 
     main = paste("Random Forest - Accuracy =",round(confMatRF$overall['Accuracy'], 4)))
```
```{r}
#Decision Trees Model Fit
set.seed(33245)
modFitDT<-rpart(classe~.,data=TrainSetS,method="class")
fancyRpartPlot(modFitDT)
```
```{r}
#Apply Model to Testdata
predictDT <- predict(modFitDT, newdata=TestSetS, type="class")
confMatDT<-confusionMatrix(predictDT,TestSetS$classe)
confMatDT
```
```{r}
#plot matrix results
plot(confMatDT$table, col = confMatDT$byClass, 
     main = paste("Decision Tree - Accuracy =",
                  round(confMatDT$overall['Accuracy'], 4)))
```
```{r}
# Generalized Boosted Model Fit
set.seed(33245)
controlGBM <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
modFitGBM  <- train(classe ~ ., data=TrainSetS, method = "gbm",trControl = controlGBM, verbose = FALSE)
modFitGBM$finalModel
```
```{r}
#prediction on Test dataset
predictGBM <- predict(modFitGBM, newdata=TestSetS)
confMatGBM <- confusionMatrix(predictGBM, TestSetS$classe)
confMatGBM
```
```{r}
#plot matrix results
plot(confMatGBM$table, col = confMatGBM$byClass, 
     main = paste("GBM - Accuracy =", round(confMatGBM$overall['Accuracy'], 4)))
```
Applying the selected Model to the Test Data
Accuracy of the 3 models are:
Random Forest=0.9998
Decision Tree=0.9998
GBM=0.9998
In this case, since all 3 models have thesame accuracy, any one of them can be selected. GBM is selected 
```{r}
predictTEST <- predict(modFitGBM, newdata=testing)
predictTEST
```